### 2016-06-26 ansi-isolation.tex

Fig.~\ref{fig:ansi-isolation} shows the specification of standard
isolation guarantees expressed as constraints over trace
well-formedness. For brevity and convenience, we adopt few notations
in this and subsequent sections.  An execution trace is destructed
into $\A$ and $\visZ$ whenever individual components of the pair are
needed. Otherwise, it is written as $\E$. Sometimes, the dot notation
(\eg~$\E.\A$) is also used. Since $\A$ and $\visZ$ are both sets, we
lift the operations on sets to pairs of sets when updating $\E$. For
example, $\E' = \E \cup (\{\eta_2\},\{(\eta_1,\eta_2)\})$ expands to
$\E' = (\E.\A \cup \{\eta_2\},\,\E.\visZ \cup \{(\eta_1,\eta_2)\})$.
When $\psi$ is a proposition, $\underE{\psi}$ denotes the
interpretation of $\psi$ in the context of the trace $\E$. Such
interpretations are defined on a case-by-case basis in
Fig.~\ref{fig:ansi-isolation}. In the following, we give informal
explanations for each definition.

In the context of a trace $(\A,\visZ)$, an effect $\eta$ is said to
belong to a transaction $T_i$ if $\eta$ belongs to the effect set $A$
and its transaction identifier is $T_i$. The containment relation is
trivially lifted to the set of effects to define $\underE{S \subseteq
T_i}$.  The interpretation of $\eta_1 \visar \eta_2$ and $\eta_1
\soar \eta_2$ are straightforward and explained in \S\ref{sec:syntax}.
A transaction $T_i$ is said to be visible to an effect $\eta$ if every
effect $\eta_1$ of $T_i$ recorded by the trace is visible to $\eta$.
$T_i$ may be visible to $\eta$ but may not be visible to every other
effect in the transaction. For a transaction $T_i$ to be considered to
be visible to a transaction $T_j$ in the context of a trace $\E$
(written $\underE{T_i \visar T_j}$), every effect ($\eta_1$) of $T_i$
present in $\E$ must be visible to every effect ($\eta_2$) of $T_j$ in
$\E$. Conversely, if none of the effects of $T_i$ present in $\E$ are
visible to any effect of $T_j$, then $T_i$ is considered invisible to
$T_j$ under $\E$ (written $\underE{T_i \invisar T_j}$). Transaction
$T_i$ is said to have written to a variable $X$ under $\E$ (i.e.,
$\underE{T_i \wrstoar X}$) if there exists a write-to-$X$ effect from
$T_i$ in $\E$. \emph{Reads-from} is similar. $T_i$ \emph{uses} $X$ if
it reads or writes $X$.

@@

In contrast to recent proposals (\emph{e.g.},
~\cite{gotsmanconcur15}), our specifications for {\sc si} and {\sc
  ser} do not necessarily impose a total order among transactions
(conflicting or otherwise). In reality, a total order under {\sc ser}
(resp. {\sc si}) is guaranteed only if the store executes all
transactions under {\sc ser} (resp. {\sc SI}) isolation. Our
specifications admit this possibility, and derive a total order under
the assumption of homogenity. However, databases almost always allow
isolation levels to be configured on a per-transaction basis, allowing
transactions at various isolation levels to coexist.  Specifications
that assume homogenity are incorrect under this setting.

>>>>

Fig.~\ref{fig:ansi-isolation} shows the specification of standard
isolation guarantees expressed as constraints over trace
well-formedness. For brevity and convenience, we adopt few notations
in this and subsequent sections.  An execution trace is destructed
into $\A$ and $\visZ$ whenever individual components of the pair are
needed. Otherwise, it is written as $\E$. Sometimes, the dot notation
(\eg~$\E.\A$) is also used. Since $\A$ and $\visZ$ are both sets, we
lift the operations on sets to pairs of sets when updating $\E$. For
example, $\E' = \E \cup (\{\eta_2\},\{(\eta_1,\eta_2)\})$ expands to
$\E' = (\E.\A \cup \{\eta_2\},\,\E.\visZ \cup \{(\eta_1,\eta_2)\})$.
When $\psi$ is a formula, $\underE{\psi}$ denotes the
interpretation of $\psi$ in the context of the trace $\E$. Such
interpretations are defined on a case-by-case basis in
Fig.~\ref{fig:ansi-isolation}. In the following, we give informal
explanations for each definition.

In the context of a trace $(\A,\visZ)$, an effect $\eta$ is said to
belong to a transaction $T_i$ if $\eta$ belongs to the effect set $A$
and its transaction identifier is $T_i$. The containment relation is
trivially lifted to the set of effects to define $\underE{S \subseteq
  T_i}$.  The interpretation of $\eta_1 \visar \eta_2$ and $\eta_1
\soar \eta_2$ are as explained earlier.  A transaction $T_i$ is said
to be visible to an effect $\eta$ if every effect $\eta_1$ of $T_i$
recorded by the trace is visible to $\eta$.  $T_i$ may be visible to
$\eta$ but may not be visible to every other effect in the
transaction. For a transaction $T_i$ to be considered to be visible to
a transaction $T_j$ in the context of a trace $\E$ (written
$\underE{T_i \visar T_j}$), every effect ($\eta_1$) of $T_i$ present
in $\E$ must be visible to every effect ($\eta_2$) of $T_j$ in
$\E$. Conversely, if none of the effects of $T_i$ present in $\E$ are
visible to any effect of $T_j$, then $T_i$ is considered invisible to
$T_j$ under $\E$ (written $\underE{T_i \invisar T_j}$). Transaction
$T_i$ is said to have written to a variable $X$ under $\E$ (i.e.,
$\underE{T_i \wrstoar X}$) if there exists a write-to-$X$ effect from
$T_i$ in $\E$. \emph{Reads-from} is similar. $T_i$ \emph{uses} $X$ if
it reads or writes $X$.

@@

%% SJ: Not sure this paragraph is necessary here.
%% In contrast to recent proposals (\emph{e.g.},
%% ~\cite{gotsmanconcur15}), our specifications for {\sc si} and {\sc
%%   ser} do not necessarily impose a total order among transactions
%% (conflicting or otherwise). In reality, a total order under {\sc ser}
%% (resp. {\sc si}) is guaranteed only if the store executes all
%% transactions under {\sc ser} (resp. {\sc si}) isolation. Our
%% specifications admit this possibility, and derive a total order under
%% the assumption of homogenity. However, databases almost always allow
%% isolation levels to be configured on a per-transaction basis, allowing
%% transactions at various isolation levels to coexist.  Specifications
%% that assume homogenity are incorrect under this setting.
